---
title: "Lecture 2 — Visualization I"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

```{r}
library(ggplot2)
```

```{r}
head(mpg)
```

```{r}
summary(mpg$hwy)
summary(mpg$cty)
```

```{r}
hist(mpg$hwy)
```

Let's use ggplot2:

```{r}
ggplot(mpg) + geom_histogram(aes(x=hwy))
```

Let's look at `hwy` and `displ`:

```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=displ))
```

Does using `cty` make a difference? This is a different measure of fuel efficiency. 

```{r}
ggplot(mpg) + geom_point(aes(x=cty, y=displ))

```

Nope, that doesn't make a difference

```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=displ, color=trans))
```

```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=displ, color=class))
```


## Fuel Efficiency By Year

```{r}
ggplot(mpg) + geom_point(aes(x=year, y=hwy))
```

```{r}
table(mpg$year)
```

```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=class, color=year))
```

```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=class, color=as.factor(year)))
```



```{r}
ggplot(mpg) + geom_boxplot(aes(y=hwy, x=class))
```


Note: If you wanted to flip the axes, use `coord_flip()`. For example:


```{r}
ggplot(mpg) + geom_boxplot(aes(y=hwy, x=class)) + coord_flip()
```

```{r}
ggplot(mpg) + geom_boxplot(aes(y=hwy, x=class)) + geom_point(aes(x=class, y=hwy, color=as.factor(year)))
```

Due to overplotting: I am thinking about using some jitter to disperse the points. 
```{r}
ggplot(mpg) + geom_boxplot(aes(y=hwy, x=class)) + geom_jitter(aes(x=class, y=hwy, color=as.factor(year)))
```

We could, if we liked this plot, customize it a bit. 


```{r}
ggplot(mpg) + geom_jitter(aes(x=class, y=hwy, color=as.factor(year))) + geom_boxplot(aes(y=hwy, x=class))
```


## Facetting 

```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=displ)) + facet_wrap(~ class)
```

```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=displ)) + facet_wrap(~ year)
```
```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=displ, color=as.factor(year)))
```

Some fun before next time:

```{r}
ggplot(mpg) + geom_point(aes(x=hwy, y=displ, color=as.factor(year))) + geom_smooth(aes(x=hwy, y=displ, color=as.factor(year)))
```

Exercises:

 - Keep exploring the `mpg` data.
 - Look up how to change the size of points.
 - Look up how to change the shape of points.
 - Make a violin plot of the hwy by class.
