---
title: "Lecture 5"
output: html_notebook
---

```{r}
library(tidyverse)
```

## Using R default methods:

```{r}
recomb <- read.table("hapmapRecombRate.bed", sep="\t")
head(recomb)
```

We can manually change the column names:

```{r}
colnames(recomb)
```

```{r}
colnames(recomb) <- c('chr', 'start', 'end', 'recomb')
head(recomb)
```

## Using the readr package

```{r}
recomb <- read_tsv('./hapmapRecombRate.bed')

head(recomb)
```

But beware! This loaded the first row as the column names.

The lesson here is to always look at the data after you read it in.

One solution is to specify `col_names=FALSE`:

```{r}
recomb <- read_tsv('./hapmapRecombRate.bed', col_names=FALSE)
```

But then we'd need to change these with `colnames()`
But, there's an even simpler way — specify the column names in the `read_tsv()` function:

```{r}
recomb <- read_tsv('./hapmapRecombRate.bed',
                   col_names=c("chr", "start" , "end", "recomb"))
```



```{r}
head(recomb)
```

Always check that the column types are as you expect!!

```{r}
ggplot(recomb) + geom_point(aes(x=start, y=recomb))
```

How many chromosomes are in the data set?

```{r}
dim(recomb)
```

```{r}
unique(recomb$chr)
```

```{r}
table(recomb$chr)
```

```{r}
ggplot(recomb, aes(x=start, y=recomb)) + geom_point() + facet_wrap(~ chr)
```

```{r}
ggplot(recomb, aes(x=start, y=recomb)) + facet_wrap(~ chr) + geom_smooth()
```


We ditched the points because outliers were driving this plot. 

## Data Loading Adventure!

```{r}
list.files()
```

```{r}
d <- read_tsv('iris.txt')
```

```{r}
as.numeric(d$Sepal.Width)
```

We see some missing values! Strange.


Goals:
 - why are some of the values above missing? Can we fix them all (or leave them missing?)
 - Count the number of unique species in the last column. What's wrong? Use the hints below to help fix the errors.
 - Plot the data. Is there anything that looks like it needs to be fixed?
 
 
```{r}
unique(d$Species)
```


Some hints: The function `which()` is very useful — read it's documentation.

To assign a new value to a column/row in R, use the follow example as an illustration:

```{r}
mistake <- data.frame(animal=c("frog", "frog", "dog", "dag", "frag"))
mistake
```

```{r}
mistake$animal[4] <- "dog"
mistake$animal[5] <- "frog"
mistake
```

