---
title: "Lecture 3 — Visualization II"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

```{r}
library(tidyverse)
```

```{r}
ggplot(diamonds) + geom_point(aes(x=carat, y=price), color='seagreen2')
```

This is different from mapping a column to a color:
```{r}
ggplot(diamonds) + geom_point(aes(x=carat, y=price, color=cut))
```

```{r}
ggplot(diamonds) + geom_point(aes(x=carat, y=price), size=0.5, alpha=1/100)
```
We're able to see the  story of the data only after dealing with overplotting. Larger diamonds are rare and expensive.



```{r}
ggplot(data=diamonds, mapping=aes(x=carat, y=price, color=cut)) + geom_point(size=0.3, alpha=1/100) + geom_smooth()
```

We can also specify the model to use, e.g. `method='loess'` or `method='lm'`. 
```{r}
ggplot(data=diamonds, mapping=aes(x=carat, y=price, color=cut)) + geom_point(size=0.3, alpha=1/100) + geom_smooth(method='loess')
```

```{r}
ggplot(mpg, aes(x=displ, y=hwy, color=class)) + geom_point() + geom_smooth()
```

Yikes! This is a great example of overfitting! See the purple line for an example.
```{r}
ggplot(mpg, aes(x=displ, y=hwy, color=class)) + geom_point() + geom_smooth(method='lm')
```
If we didn't want multiple lines, provide different `aes()`:

```{r}
ggplot(mpg) + geom_point(aes(x=displ, y=hwy, color=class)) + geom_smooth(aes(x=displ, y=hwy), method='lm')
```


## Refining Plots:

Labels!

```{r}
ggplot(faithful) + geom_point(aes(eruptions, waiting)) + ggtitle("Old Faithful") + xlab("Eruptions (minutes)")+ ylab("Waiting (minutes)")
```

```{r}
ggplot(faithful) + geom_point(aes(eruptions, waiting)) + ggtitle("Old Faithful") + xlab("Eruptions (minutes)")+ ylab("Waiting (minutes)") + theme_bw()
```
```{r}
ggplot(faithful) + geom_point(aes(eruptions, waiting)) + ggtitle("Old Faithful") + xlab("Eruptions (minutes)")+ ylab("Waiting (minutes)") + theme_classic()
```
```{r}
library(ggthemes)
ggplot(faithful) + geom_point(aes(eruptions, waiting)) + ggtitle("Old Faithful") + xlab("Eruptions (minutes)")+ ylab("Waiting (minutes)") + theme_economist()
```
```{r}
ggplot(faithful) + geom_point(aes(eruptions, waiting)) + ggtitle("Old Faithful") + xlab("Eruptions (minutes)")+ ylab("Waiting (minutes)") + theme_tufte()
```
```{r}
ggplot(faithful) + geom_point(aes(eruptions, waiting)) + ggtitle("Old Faithful") + xlab("Eruptions (minutes)")+ ylab("Waiting (minutes)") + theme_fivethirtyeight()
```

Other customizations are possible! See the ggplot2 documentation.

Also! You can use R base graphics, e.g.

```{r}
plot(faithful$eruptions, faithful$waiting)
```

## Geom Line and Dataframes

```{r}
ex_data <- data.frame(timepoint=c(1, 2, 3), value=c(1, 6, 7))
ex_data
```

```{r}
ggplot(ex_data) + geom_line(aes(x=timepoint, y=value)) + geom_point(aes(x=timepoint, y=value))
```

Now, let's try something — https://en.wikipedia.org/wiki/Don%27t_repeat_yourself
```{r}
ggplot(ex_data, aes(x=timepoint, y=value)) + geom_point(size=2) + geom_line(linetype=2) 
```

## Last Bit: Density and Frequency Plots

```{r}
ggplot(diamonds)+ geom_freqpoly(aes(x=carat, color=cut))
```

```{r}
ggplot(diamonds)+ geom_point(aes(x=carat, y=price, color=cut))  + facet_wrap(~clarity)
```